---
title: "Developing Data Products Course Project"
author: "Davy Meesemaecker"
date: "15/2/2018"
output:
  ioslides_presentation: default
  slidy_presentation: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Introduction

This peer assessed assignment has two parts :

- Create a Shiny application and deploy it on Rstudio's servers. 
- Use Slidify or Rstudio Presenter to prepare a reproducible pitch presentation about the application.

## The application

For the application, I used a population dataset from french population in 2009 available here
<a href="http://sql.sh/736-base-donnees-villes-francaises">Sql.sh</a>. In it, we'll find the population for around 36700 cities (called communes in french).

What can you do with the application ?

- The left side with two different sliders in the grey area. First slider allows you to define the range of population you'd like to see. This range goes from 0 to 2 300 000. The second slider lets you choose the number of breaks in the histogram with a range from 10 to 100.

- The right side with a table on top listing all cities included in the range you previously selected and a histogram below gives you the estimate of the population range distribution.

## The data

Let's take a short look at the code and data before it is transformed into an application :

```{r echo=TRUE, message=FALSE, warning=FALSE}
library(dplyr)
villes <- read.csv("villes_france.csv")
villes <- villes[, c("Ozan", "X618", "X4.91667", "X46.3833")]
colnames(villes) <- c("Ville", "Population", "lng", "lat")
villes <- arrange(villes, desc(Population))
head(villes)
```


## Where can you find the application ?

The application can be found here :

<a href="https://dasmey.shinyapps.io/final_project/">Application</a>

You can also find the code here :

<a href="https://github.com/dasmey/Developing-Data-Products/tree/gh-pages/final_project">Code</a>


