---
title: "Regression Models Course Project"
author: "Davy Meesemaecker"
date: "21/1/2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Executive summary

You work for Motor Trend, a magazine about the automobile industry. Looking at a data set of a collection of cars, they are interested in exploring the relationship between a set of variables and miles per gallon (MPG) (outcome). They are particularly interested in the following two questions:

° “Is an automatic or manual transmission better for MPG”

° "Quantify the MPG difference between automatic and manual transmissions"

In the report, we fit three different models. The first model is a simple model with only one variable (auto/manual), which suggests that manual transmission’s MPG is 7.245 miles higher than auto transmission, but the model only explained 33.85% of the variance in MPG. The second model is using all variables does not have any coefficients which appear to be significant. The last model, a model with 4 variables (cylinder, horsepower, weight and auto/manual) which explains 86% of the variance in MPG, suggests that manual transmission’s MPG is still higher than auto transmission’s MPG, even though the miles per gallon saved is only higher by 1.81 miles compared to auto transmission after including the three other variables - cylinder, horsepower and weight.

# Data analysis

The data was extracted from the 1974 Motor Trend US magazine, and comprises fuel consumption and 10 aspects of automobile design and performance for 32 automobiles (1973–74 models).

```{r message=FALSE, warning=FALSE}
library(ggplot2)
library (dplyr)
data("mtcars")
str(mtcars)
sum(is.na(mtcars))
```

By taking a quick look at the data, we learn our data doesn't contain any NA and that all variables are numeric for now but some of them need be converted to factors

```{r}
mtcars$cyl <- as.factor(mtcars$cyl)
mtcars$vs <- as.factor(mtcars$vs)
mtcars$am <- as.factor(mtcars$am)
mtcars$gear <- as.factor(mtcars$gear)
mtcars$carb <- as.factor(mtcars$carb)
```

Now it's done, we can get a first look on how transmission type impacts the fuel consumption

```{r}
library(plyr)
mtcars$transmission <- revalue(mtcars$am, c("0" = "automatic", "1" = "manual"))
mtcars <- mtcars[, -c(9)]
ggplot(mtcars, aes(x = transmission, y = mpg, fill = transmission)) + geom_boxplot() + geom_jitter(colour = "darkblue", alpha=0.6) + labs(x = "Type of transmission", y = "Miles per Gallon")
```

The plot suggests an impact of the type of transmission on the fuel consumption as Automatic cars seem to be less fuel efficient than manual cars. Fitting a simple model could help us validate this trend.

```{r}
mdl <- lm(mpg ~ transmission, data = mtcars)
summary(mdl)
```

So what do we learn ? First, our manuel transmission estimate is 7.245, implying a manual car can drive 7.245 miles per gallon more than an automatic car on average. It's approuved by our very low p-value, making our model is significant at a 0.001 level. However, we only cover 35% of mpg's variability and it seems obvious as we only had one predictor in our model. What if we add all variables in our model ?

```{r}
mdl2 <- lm(mpg ~. , data = mtcars)
summary(mdl2)
```

Now we cover 89% of mpg's variability but unfortunately, none of the predictors is significant at 0.05. We need to find a better model by removing some noisy variables.

```{r eval=FALSE}
bestformula <- step(mdl2)
```

Now we can fit our final model

```{r}
bestmdl <- lm(mpg ~ cyl + hp + wt + transmission, data = mtcars)
anova(mdl2, bestmdl)
summary(bestmdl)
```

So this model covers 86% of the variance of mpg with four of all predictors, cyl(number of cylinders), hp(horsepower), weight(of the car) and transmission type. it obviously covers less than our model with all the predictors but it covers more than enough variability. Most of all, Our model p-value is very low, 1.506e-10, meaning our final model is reliable at more than 99%.

```{r}
par(mfrow = c(2,2))
plot(bestmdl)
```

# Conclusion :

- Is an automatic or manual transmission better for MPG?

In our final model, It appears that manual transmission cars are better for MPG compared to automatic cars but less than in our very first model comparing mpg and transmission. Other variables seem more reliable to explain mpg's variability like the number of cylinders, the gross horsepower and the weight of the car.

- Quantify the MPG difference between automatic and manual transmissions :

At first sight, when simply comparing miles per gallon(mpg) to the transmission type, it seems like manual transmission was more efficient for mpg with 7.245 miles more than automatic cars on average. When comparing mpg with the number of cylinders, the gross horspower and the weight, it drops to 1.81.